Online Multiclass Linear Classification with Bandit Feedback

### 2019-01-14 ICML-submission/algorithm-for-strongly-linearly-separable-data.tex
Fix introduction

Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}
can be extended to nonlinear classification using kernels~\cite{Scholkopf-Smola-Bach-2002}.
For each class, as opposed
to explicitly maintaining the predictor's weights for each class, we maintain
the set of examples whose scaled feature transformations are added to the weights in Perceptron updates.
We give a formal description of the kernelized algorithm in Algorithm~\ref{algorithm:kernelized}.

>>>>

Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples}
can be extended to nonlinear classification using kernels~\cite{Scholkopf-Smola-2002, Shawe-Taylor-Cristianini-2004}.
For each class, as opposed
to explicitly maintaining the predictor's weights for each class, we maintain
the set of examples whose scaled feature transformations are added to the weights in Perceptron updates.
We give a formal description of the kernelized algorithm in Algorithm~\ref{algorithm:kernelized}.

### 2019-01-14 ICML-submission/introduction.tex
Fix introduction

\begin{protocol}[h]
\caption{\textsc{Online Multiclass Classification with Bandit Feedback}
\label{algorithm:game-protocol}}
\begin{algorithmic}[1]
{
\REQUIRE{Number of classes $K$, number of rounds $T$.}
%\REQUIRE{.}
\FOR{$t=1,2,\dots,T$}
\STATE Adversary chooses example $(x_t, y_t)$, where $x_t$ is revealed to the learner.
%\STATE{Observe feature vector $x_t$}
%\STATE{Adversary \emph{secretly} chooses correct class label $y_t \in \{1,2,\dots,K\}$}
%\chicheng{Or we simply say: Adversary chooses example $(x_t, y_t)$, where $x_t$ is revealed to the learner.}
\STATE{Predict class label $\widehat y_t \in \{1,2,\dots,K\}$}
\STATE{Observe feedback $z_t \in \{0,1\}$ \\ \qquad where $z_t = \begin{cases}
0 & \text{if $\widehat y_t = y_t$,} \\
1 & \text{if $\widehat y_t \neq y_t$.}
\end{cases}$
}
\ENDFOR
}
\end{algorithmic}
\end{protocol}

The performance is of the learner is measured by its cumulative number of
mistakes $\sum_{t=1}^T z_t = \sum_{t=1}^T \indicator{\widehat y_t \neq y_t}$.
The bandit feedback is a natural feedback model motivated by applications
including online advertising and online content optimization.

In this paper, we focus on the special case when the examples chosen by the
adversary lie in $\R^d$ and are linearly separable with a margin. We introduce
two notions of linear separability, \emph{weak} and \emph{strong}. These are
formally stated in
Definitions~\ref{definition:weak-linear-separability}~and~\ref{definition:strong-linear-separability}.
The standard notion of multiclass linear separability corresponds to the weak
linear separability. For binary classification, these two notions are
identical. However, they differ for multiclass classification with $K \geq 3$.
For multiclass classification with $K$ classes, the standard notion of linear
separability means that the examples from each class lie in an intersection of
$K-1$ halfspaces and the examples outside of the class lie in the complement of
the intersection of the halfspaces. Strong linear separability means that
examples from each class are separated from the remaining examples by a
\emph{single} hyperplane.

@@

The bandit feedback setting, however, is much more challenging. For the case
when the examples are strongly linearly separable, we design a simple and
efficient algorithm with an expected number of mistakes at most $(K-1) \lfloor
(R/\gamma)^2 \rfloor$. The algorithm can be viewed as running $K$ copies of the
\textsc{Binary Perceptron} algorithm, one copy for each class. Intuitively
speaking, the factor $K-1$ is the price we pay for the bandit feedback, or more
precisely, the lack of full-information feedback. We also prove that any
(possibly randomized) algorithm makes $\Omega(K (R/\gamma)^2))$ mistakes in the
worst case.

For the standard notion of linear separability, we propose several algorithms
with various time-complexity-versus-mistake-bound trade-offs.
\chicheng{I wonder if we should just present the results using the rational kernel,
 as results using other kernels are nonadaptive,
 and achieve qualitatively similar mistake bounds.}
First class of
algorithms map the examples into a high-dimensional inner product space. In this
space the examples become \emph{strongly} linearly separable. This allows us to
run the algorithm for the strongly linearly separable case. All of these
algorithms can be kernelized. (Some of them have to be kernelized since the
underlying inner product space is infinite dimensional.) Kernelization does not
change the input-output behavior of the algorithm, it only affects its time
and memory complexity.

These algorithms differ by choice of the mapping or, equivalently, choice of the
kernel. The crux of the problem is the analysis of the approximation properties
of these mappings/kernels and the way margin in the original space $\R^d$ under
standard linear separability notion gets transformed into a margin in the
high-dimensional space under the strong separability notion. This problem is
related to the problem of learning intersection of halfspaces and has been
studied previously by \cite{Klivans-Servedio-2008}. As an important technical
achievement, we improve on the results of \cite{Klivans-Servedio-2008} by
removing completely the dependency on the original dimension $d$.

Finally, we propose a very different algorithm that is based on the obvious idea
that two points that are close enough must have the same label. Here, close
enough means that they cannot be separated from each other by a linear separator
with a margin.

All the algorithms run in time that is polynomial in the dimension of the
feature vectors, the number of classes and the number of rounds. At the same
time, all the algorithms make number of mistakes that does \emph{not} depend
on the number of rounds. Instead, their number of mistakes depends only on the
margin, number of classes, dimension of the feature vectors, and the radius of
the ball in which the data lie.

Finally, we study two questions related to the computational and
information-theoretic hardness of the problem. Any algorithm for the bandit
setting collects information in the form of so called \emph{strongly labeled}
and \emph{weakly labeled} examples. Strongly labeled examples are those for
which we know the class label. Weakly labeled example is an example for which we
know that class label can be anything except for a particular one class. First,
we show the offline problem of finding a linear multiclass separator consistent
with a set of strongly and weakly labeled examples is NP-hard. This result is a
consequence of the fact that finding an intersection of two halfspaces
consistent with a set of binary-labeled examples is
NP-hard~\citep{Blum-Rivest-1993}. Second, we show that any algorithm that uses
only strongly-labeled examples and ignores weakly labeled examples makes at
least $\Omega(???)$ mistakes.

>>>>

\begin{protocol}[h]
\caption{\textsc{Online Multiclass Classification with Bandit Feedback}
\label{algorithm:game-protocol}}
\begin{algorithmic}[1]
{
\REQUIRE{Number of classes $K$, number of rounds $T$.}
\FOR{$t=1,2,\dots,T$}
\STATE Adversary chooses example $(x_t, y_t)$, where $x_t$ is revealed to the learner.
\STATE{Predict class label $\widehat y_t \in \{1,2,\dots,K\}$}
\STATE{Observe feedback $z_t \in \{0,1\}$ \\ \qquad where $z_t = \begin{cases}
0 & \text{if $\widehat y_t = y_t$,} \\
1 & \text{if $\widehat y_t \neq y_t$.}
\end{cases}$
}
\ENDFOR
}
\end{algorithmic}
\end{protocol}

The performance of the learner is measured by its cumulative number of
mistakes $\sum_{t=1}^T z_t = \sum_{t=1}^T \indicator{\widehat y_t \neq y_t}$.
The bandit feedback is a natural feedback model motivated by applications
including online advertising and online content optimization.

In this paper, we focus on the special case when the examples chosen by the
adversary lie in $\R^d$ and are linearly separable with a margin. We introduce
two notions of linear separability, \emph{weak} and \emph{strong}. These are
formally stated in
Definitions~\ref{definition:weak-linear-separability}~and~\ref{definition:strong-linear-separability}.
The standard notion of multiclass linear separability corresponds to the weak
linear separability. For binary classification, the weak and strong separability
are identical. However, they differ for multiclass classification with $K \ge 3$
classes. For multiclass classification with $K$ classes, the standard notion of
linear separability means that the examples from each class lie in an
intersection of $K-1$ halfspaces and the examples outside of the class lie in
the complement of the intersection of the halfspaces. Strong linear separability
means that examples from each class are separated from the remaining examples by
a \emph{single} hyperplane.

@@

The bandit feedback setting, however, is much more challenging. For the case
when the examples are strongly linearly separable, we design a simple and
efficient algorithm with an expected number of mistakes at most $(K-1) \lfloor
4(R/\gamma)^2 \rfloor$. The algorithm can be viewed as running $K$ copies of the
\textsc{Binary Perceptron} algorithm, one copy for each class. Intuitively
speaking, the factor $K-1$ is the price we pay for the bandit feedback, or more
precisely, the lack of full-information feedback. We also prove that any
(possibly randomized) algorithm makes $\Omega(K (R/\gamma)^2))$ mistakes in the
worst case.

For the case when examples are weakly linearly separable, we construct a feature
mapping $\phi$ such that makes them \emph{strongly} linearly separable. We then
use the kernelized version of the algorithm for the strongly separable case; for
more details on kernel methods see e.g.~\citet{Scholkopf-Smola-2002} and
\citet{Shawe-Taylor-Cristianini-2004}. The kernel $k(x,y)$ corresponding to the
feature mapping $\phi$ has a simple explicit formula and can be computed in
$O(d)$ time.

The number of mistakes of the kernelized algorithm depends on the margin in the
corresponding feature space. We analyze how the margin parameter of weak
separability in the original space $\R^d$ gets transformed into a margin
parameter of strong separability in the new feature space. This problem is
related to the problem of learning intersection of halfspaces and has been
studied previously by \citet{Klivans-Servedio-2008}. In fact, our analysis can
be viewed as a follow up on their work. We improve on the results of
\citet{Klivans-Servedio-2008} by removing the dependency on the original
dimension $d$.

The resulting kernelized algorithm runs in time that is polynomial in the
original dimension of the feature vectors $d$, the number of classes $K$, and
the number of rounds $T$. We prove that if the examples lie in unit ball of
$\R^d$ and are weakly separable with margin $\gamma$, the algorithm makes at
most $O(???)$ mistakes.

### 2019-01-14 ICML-submission/notions-of-linear-separability.tex
Fix introduction

The notion of strong linear separability has appeared in the literature
(see e.g.~\cite{Chen-Chen-Zhang-Chen-Zhang-2009}).
\footnote{Although~\cite{Chen-Chen-Zhang-Chen-Zhang-2009} claimed that their
Conservative OVA algorithm with PA-I update has a finite mistake bound under the strong
linear separability assumption, their proof of Theorem 2 is problematic: their
Lemma 4 cannot be directly applied to the bandit setting,
as the algorithm only makes updates on a subset of the $w_i$'s.}
Intuitively, strong linear
separability means that, for each class $i$, the set of examples belonging to class $i$ and
the set of example belonging to the remaining $K-1$ classes are separated by
a linear classifier $w_i$ by a margin $\frac \gamma 2$.
%Formally, $w_i$ separates the set of examples $\cbr{(x_t, y_t): t \in
%\cbr{1,\ldots,T}, y_t = i}$ and $\cbr{(x_t, y_t): t \in \cbr{1,\ldots,T}, y_t
%\neq i}$.

>>>>

The notion of strong linear separability has appeared in the literature;
see e.g.~\citet{Chen-Chen-Zhang-Chen-Zhang-2009}.
\footnote{Although~\cite{Chen-Chen-Zhang-Chen-Zhang-2009} claimed that their
Conservative OVA algorithm with PA-I update has a finite mistake bound under the strong
linear separability assumption, their proof of Theorem 2 is incorrect. Their
Lemma 4 cannot be directly applied to the bandit setting,
as the algorithm only makes updates on a subset of the $w_i$'s.}
Intuitively, strong linear
separability means that, for each class $i$, the set of examples belonging to class $i$ and
the set of example belonging to the remaining $K-1$ classes are separated by
a linear classifier $w_i$ by a margin $\frac \gamma 2$.
%Formally, $w_i$ separates the set of examples $\cbr{(x_t, y_t): t \in
%\cbr{1,\ldots,T}, y_t = i}$ and $\cbr{(x_t, y_t): t \in \cbr{1,\ldots,T}, y_t
%\neq i}$.

### 2019-01-14 ICML-submission/title-and-abstract.tex
Fix introduction

\item Under the more challenging weak linear separability condition, we design
an efficient algorithm with a mistake bound of $2^{\widetilde{O}(\min(K \log^2
\frac{1}{\gamma}, \log K \sqrt{\frac 1 \gamma}))}$ \footnote{We use the notation
$\widetilde{O}(f(\cdot)) = O(f(\cdot) \polylog(f(\cdot)))$.}. Our algorithm is
based on a infinite-dimensional feature mapping that transforms weak linear
separable examples to strong linear separable examples, which in turn is
inspired by the work of \citet{Klivans-Servedio-2008} on learning intersection of
halfspaces.
\end{enumerate}
\end{abstract}

>>>>

\item Under the more challenging weak linear separability condition, we design
an efficient algorithm with a mistake bound of $2^{\widetilde{O}(\min(K \log^2
\frac{1}{\gamma}, \log K \sqrt{\frac 1 \gamma}))}$ \footnote{We use the notation
$\widetilde{O}(f(\cdot)) = O(f(\cdot) \polylog(f(\cdot)))$.}. Our algorithm is
based on a infinite-dimensional feature mapping that transforms weakly linear
separable examples to strongly linear separable examples, which in turn is
inspired by the work of \citet{Klivans-Servedio-2008} on learning intersection of
halfspaces.
\end{enumerate}
\end{abstract}
